On the Limitations of Weighted Model Counting

### 2020-03-28 paper.tex
One more lemma

For clarity and succinctness, we will occasionally use three other operations
that can be defined using the original three\footnote{We use $+$ to denote
  symmetric difference because it is the additive operation of a Boolean ring.}:
\begin{align*}
  a \to b &= \neg a \lor b, \\
  a \leftrightarrow b &= (a \land b) \lor (\neg a \land \neg b), \\
  a + b &= (a \land \neg b) \lor (\neg a \land b).
\end{align*}
We can also define a partial order $\le$ on $\mathbf{B}$ as $a \le b$ if $a = b
\land a$ (or, equivalently, $a \lor b = b$) for $a, b \in \mathbf{B}$.
Furthermore, let $a < b$ denote $a \le b$ and $a \ne b$. For the rest of this
paper, let $\mathbf{B}$ refer to the BA $(\mathbf{B}, \land, \lor, \neg, 0, 1)$.
For any $S \subseteq \mathbf{B}$, we write $\bigvee S$ for $\bigvee_{x \in S} x$
and call it the \emph{supremum} of $S$. Similarly, $\bigwedge S = \bigwedge_{x
  \in S} x$ is the \emph{infimum}. By convention, $\bigwedge \emptyset = 1$ and
$\bigvee \emptyset = 0$.

@@

@@

\begin{corollary}
  Let $I \subseteq \mathbf{B}$ be an ideal. If $\mathbf{B}$ is atomic, then so
  is $\mathbf{B}/I$.
\end{corollary}

\begin{lemma}
  Let $(p)$ be a principal ideal. For any $a \in \mathbf{B}$, if $a/(p)$ is an
  atom in $\mathbf{B}/(p)$, then $a \land \neg p$ is an atom in $\mathbf{B}$
  such that $(a \land \neg p)/I = a/I$, and $m^*(a/(p)) = m(a \land \neg p)$.
\end{lemma}

\begin{theorem}[Using the previous three results]
  Let $L \subset \mathbf{B}$ and $w\colon L \to \mathbb{R}_{\ge 0}$ be such that
  $\WMC_w\colon \mathbf{B} \to \mathbb{R}_{\ge0}$ is a WMC measure. Let $p \in
  \mathbf{B}$ be such that $\WMC(p) = 0$, and let $g\colon \mathbf{B}/(p) \to
  \mathbf{B}$ be defined as $g(a/(p)) = a \land \neg p$. Then $\WMC \circ g$
  is a WMC measure on $\mathbf{B}/(p)$ with respect to the same $L$ and $w$
  such that
  \[
    \begin{tikzcd}
      \mathbf{B} \arrow{d}[swap]{\WMC} & \mathbf{B}/(p) \arrow{l}[swap]{g}
      \arrow{dl}{\WMC \circ g} \\
      \mathbb{R}_{\ge 0}
    \end{tikzcd}
  \]
  commutes.
\end{theorem}

>>>>

For clarity and succinctness, we will occasionally use three other operations
that can be defined using the original three\footnote{We use $+$ to denote
  symmetric difference because it is the additive operation of a Boolean ring.}:
\begin{align*}
  a \to b &= \neg a \lor b, \\
  a \leftrightarrow b &= (a \land b) \lor (\neg a \land \neg b), \\
  a + b &= (a \land \neg b) \lor (\neg a \land b).
\end{align*}
We can also define a partial order $\le$ on $\mathbf{B}$ as $a \le b$ if $a = b
\land a$ (or, equivalently, $a \lor b = b$) for $a, b \in \mathbf{B}$.
Furthermore, let $a < b$ denote $a \le b$ and $a \ne b$. For the rest of this
paper, let $\mathbf{B}$ refer to the BA $(\mathbf{B}, \land, \lor, \neg, 0, 1)$.
For any $S \subseteq \mathbf{B}$, we write $\bigvee S$ for $\bigvee_{x \in S} x$
and call it the \emph{supremum} of $S$. Similarly, $\bigwedge S = \bigwedge_{x
  \in S} x$ is the \emph{infimum}. By convention, $\bigwedge \emptyset = 1$ and
$\bigvee \emptyset = 0$. For any $a, b \in \mathbf{B}$, we say that $a$ and $b$
are \emph{disjoint} if $a \land b = 0$.

@@

\begin{lemma}[\cite{givant2008introduction}] \label{lemma:measure_and_order}
  Let $m\colon \mathbf{B} \to \mathbb{R}_{\ge 0}$ be a measure. Then for all $a,
  b \in \mathbf{B}$, if $a \le b$, then $m(a) \le m(b)$.
\end{lemma}

@@

% TODO: do I need this?
%\begin{corollary}
%  Let $I \subseteq \mathbf{B}$ be an ideal. If $\mathbf{B}$ is atomic, then so
%  is $\mathbf{B}/I$.
%\end{corollary}

\begin{lemma}
  Let $m\colon \mathbf{B} \to \mathbb{R}_{\ge 0}$ be a measure, let $p \in
  \mathbf{B}$ be such that $m(p) = 0$, and let $m^*\colon \mathbf{B}/(p) \to
  \mathbb{R}_{\ge 0}$ be a measure defined by $m^*(a/(p)) = m(a \lor p)$. Then
  for any $a \in \mathbf{B}$, if $a/(p)$ is an atom in $\mathbf{B}/(p)$, then $a
  \land \neg p$ is an atom in $\mathbf{B}$ such that $m^*(a/(p)) = m(a \land
  \neg p)$.
\end{lemma}
\begin{proof} % TODO: This proof is a mess and needs more citations.
  First, we want to show that if $a/(p)$ is an atom, then $a \land \neg p$ is an
  atom. We can instead prove the contrapositive statement, i.e., if there exists
  a $b \in \mathbf{B}$ such that $0 < b < a \land \neg p$, then there exists a
  $b' \in \mathbf{B}$ such that $0/(p) < b'/(p) < a/(p)$. We will show that, in
  fact, we set $b' = b$.
  \Cref{lemma:canonical_homomorphism,lemma:homomorphisms_and_order} already tell
  us that $b/(p) \le a/(p)$, so we only need to show that $0/(p) < b/(p) \ne
  a/(p)$. For the first part, note that
  \[
    0/(p) < b/(p) \quad \iff \quad b/(p) \ne 0/(p) \quad \iff \quad b \not\in
    (p) \quad \iff \quad b \not\le p \quad \iff \quad b \land \neg p \ne 0
  \]
  by \cref{lemma:order}. But if $b \land \neg p = 0$, then $b \land a \land \neg
  p = 0$. This contradicts either that $b \le a \land \neg p$ (i.e., $b \land a
  \land \neg p = b$) or that $b \ne 0$. For the second part, i.e., $b/(p) \ne
  a/(p)$, we will show that if $b/(p) = a/(p)$, and $b \le a \land \neg p$, then
  $b = a \land \neg p$. Indeed,
  \[
    b/(p) = a/(p) \quad \iff \quad a + b \in (p) \quad \iff \quad a+b \le p
    \quad \iff \quad (a+b) \land \neg p = 0,
  \]
  and
  \[
    (a+b) \land \neg p = [(a \land \neg b) \lor (\neg a \land b)] \land \neg p =
    (a \land \neg b \land \neg p) \lor (\neg a \land b \land \neg p),
  \]
  so $(a+b) \land \neg p = 0$ implies that $a \land \neg b \land \neg p = 0$
  which is equivalent to $a \land \neg p \le b$. Therefore we have that $a \land
  \neg p \le b \le a \land \neg p$, so $b = a \land \neg p$ which, by
  contradiction, shows that $b/(p) \ne a/(p)$ and finishes the proof that $0/(p)
  < b/(p) < a/(p)$.

  In order to show that $m^*(a/(p)) = m(a \land \neg p)$, note that $a \land p$,
  $a \land \neg p$, and $\neg a \land p$ are pairwise disjoint and their
  supremum is $a \lor p$, so we have that
  \[
    m^*(a/(p)) = m(a \lor p) =  m(a \land p) + m(a \land \neg p) + m(\neg a
    \land p).
  \]
  Furthermore, since $a \land p \le p$, $m(a \land p) \le m(p) = 0$. Similarly,
  $m(\neg a \land p) = 0$, so $m^*(a/(p)) = m(a \land \neg p)$ as required.
\end{proof}
% TODO: the undergraduate book vaguely mentions that, when taking a quotient
% w.r.t. an ideal of elements with measure zero, all elements in an equivalence
% class will have the same measure.

\begin{theorem}[Using the previous results]
  Let $L \subset \mathbf{B}$ and $w\colon L \to \mathbb{R}_{\ge 0}$ be such that
  $\WMC_w\colon \mathbf{B} \to \mathbb{R}_{\ge0}$ is a WMC measure. Let $p \in
  \mathbf{B}$ be such that $\WMC(p) = 0$, and let $g\colon \mathbf{B}/(p) \to
  \mathbf{B}$ be defined as $g(a/(p)) = a \land \neg p$. Then $\WMC \circ g$
  is a WMC measure on $\mathbf{B}/(p)$ with respect to the same $L$ and $w$
  such that
  \[
    \begin{tikzcd}
      \mathbf{B} \arrow{d}[swap]{\WMC} & \mathbf{B}/(p) \arrow{l}[swap]{g}
      \arrow{dl}{\WMC \circ g} \\
      \mathbb{R}_{\ge 0}
    \end{tikzcd}
  \]
  commutes.
\end{theorem}
